---
title: "QFunC Mixed Model"
subtitle: "Simulation results"
format: revealjs
self-contained: true
---

```{r}
#| include=FALSE, echo=FALSE, message=FALSE, warning=FALSE
library(tidyverse)
library(reshape2)
library(stringr)
library(kableExtra)
here::i_am("full-run/full-run.qmd")
library(here)
```


## Model
$J$ regions, $L_j$ voxels in region $j$, $M$ time points.

$$
X_{jlm} = \mu_j + \eta_{jm} + \gamma_{j\ell m} + \epsilon_{j\ell m}
$$

For a region $j$,

* $\mu_j$ is the overall mean
* $\eta_{jm}$ is the common mean-zero signal shared by all voxels
* $\gamma_{j\ell m}$ is the spatiotemporal signal across voxels
* $\epsilon_{j\ell m}$ is i.i.d. noise

##
<div style="font-size: 24px;">
$$
X_{j\ell m} = \mu_j + \eta_{jm} + \gamma_{j\ell m} + \epsilon_{j\ell m}
$$

\begin{align}
   \label{eq:SpaTempKern}
   (C_j)_{ll'} &= K(\big\lVert{v_{j\ell} - v_{j\ell'}}\big\rVert; \nu_j, \phi_j) \nonumber \\
   (B_j)_{mm'} &= k_{\gamma_j}H(|m - m'|; \tau_{\gamma_j}) +  \sigma^2_{\gamma_j}\delta_{mm'} \nonumber \\
   (A)_{mm'} &= H(|m - m'|; \tau_\eta) + \sigma^2_\eta \delta_{mm'} \\
\end{align}
</div>

. . .

<div style="font-size: 24px;">

\begin{align}
   \quad R &= \begin{pmatrix}
      1 & \rho_{12} & \cdots & \rho_{1J} \\
      \rho_{12} & 1 & \ddots & \cdots \\
      \vdots & \ddots & \ddots & \ddots \\
      \rho_{J1} & \vdots & \ddots & 1
   \end{pmatrix},
   \Omega = SRS, \quad S = \operatorname{diag}(k_{\eta_1}^{1/2}, \dotsc, k_{\eta_J}^{1/2}) \\
\label{eq:varMats}
    \operatorname{Var}(\epsilon) &= \Sigma_\epsilon :=
      \operatorname{diag}(\sigma^2_{\epsilon_1} I_{ML_1},\dotsc , \sigma^2_{\epsilon_J} I_{ML_J}) \\
    \operatorname{Var}(\gamma) &= \Sigma_\epsilon^{1/2}\Lambda \Sigma_\epsilon^{1/2},
    \quad \Lambda = \operatorname{diag}(C_1 \otimes B_1,\dotsc, C_J \otimes B_J) \\
    \operatorname{Var}(\eta) &= (\tilde{\Sigma}_\epsilon^{1/2}\Omega \tilde{\Sigma}_\epsilon^{1/2})\otimes A
\end{align}
</div>

. . .

<div style="font-size: 24px;">
Stage 1 parameters:
$(k_{\gamma_j}, \sigma^2_{\gamma_j}, \tau_{\gamma_j}, \phi_j, \sigma^2_{\epsilon_j})$

Stage 2 parameters:
$(\rho_{jj'}, k_{\eta_j}, k_{\eta_{j'}}, \tau_\eta, \sigma^2_{\eta})$

Need to choose suitable parameters for our simulations.
</div>

## Simulation setting

We consider various simulation settings to evaluate the performance of our method.

* Correct specification
* Spatial and temporal misspecification

For each simulation setting, we generate data from three regions using real
voxel coordinates and $60$ time points.

The three regions contain $41$, $25$, and $77$ voxels.

##

```{r}
#| echo=FALSE, message=FALSE, warning=FALSE, fig.width=12, fig.height=8, fig.align='center',
#| fig.cap="Correct specification"
create_boxplot <- function(ggdf, title, n = 100) {
   p <- ggplot(ggdf) +
   geom_boxplot(mapping = aes(x =  pair, y = value, fill = method)) +
   geom_segment(aes(x = as.numeric(pair)-0.5,
                     xend = as.numeric(pair)+0.5,
                     y = yintercept,
                     yend = yintercept),
                  lty = 2) +
   ylim(-1, 1) +
   facet_grid(delta ~ psi,
               labeller = label_bquote(rows = delta == .(delta),
                                       cols = psi == .(psi))) +
   theme_bw() +
   theme(strip.text.y = element_text(angle = 0)) +
   labs(x = "Region pair", y = "\u03C1") +
   ggtitle(title, subtitle = str_glue("n = {n}"))
   p
}
ggdf <- readRDS(here("full-run", "full-std-df.rds"))
p <- create_boxplot(ggdf, "Simulation results - correctly specified covariance structure")
print(p)
```

##
```{r}
library(dplyr)
library(kableExtra)
tabs1 <- filter(ggdf, method == "ReML") |>
         group_by(delta, psi, pair) |>
         summarize(mean = mean(value),
                   median = median(value),
                   mad = mean(abs(value - yintercept)),
                   rmse = sqrt(mean((value - yintercept)^2)),
                   sd = sd(value),
                   .groups = 'drop')

# Split the data into smaller parts
tabs1_r12 <- filter(tabs1, pair == "r12") |>
   select(-pair) |>
   mutate(across(-c(delta, psi), round, 3))
tabs1_r13 <- filter(tabs1, pair == "r13") |>
   select(-pair) |>
   mutate(across(-c(delta, psi), round, 3))
tabs1_r23 <- filter(tabs1, pair == "r23") |>
   select(-pair) |>
   mutate(across(-c(delta, psi), round, 3))
```

::: {.panel-tabset}

### Region 1 and 2

```{r}
kable(tabs1_r12, "html") |>
   kable_styling(font_size = 30)
```

### Region 1 and 3

```{r}
kable(tabs1_r13, "html") |> kable_styling(font_size = 30)
```

### Region 2 and 3

```{r}
kable(tabs1_r23, "html") |> kable_styling(font_size = 30)
```

:::

##

<div style="font-size: 24px;">
The average of intra-regional correlations is given by
$$
\alpha_j = \frac{1}{L_j^2} \sum_{\ell, \ell'} \operatorname{Cor}(Y_j(v_{j\ell, t_m}), Y_j(v_{j\ell'}, t_m))
$$
We can write
$$
\alpha_j = \delta_j + (1-\delta_j)\psi_j
$$
where

* $\delta_j = \frac{k_{\eta_j}(1+\sigma^2_\eta)}{k_{\eta_j}(1+\sigma^2_\eta) + (k_{\gamma_j} + \sigma^2_{\gamma_j})}$ is the variance of $\eta_j$ relative to $Y_j(v, t)$
* $\psi_j = \frac{1}{L_j^2} \sum_{\ell, \ell'} K(\big\lVert{v_{j\ell} - v_{j\ell'}}\big\rVert; \nu_j, \phi_j)$ is the average intra-regional spatial covariance
</div>

. . .

<div style="font-size: 24px;">

* Large $\delta_j$ means strong common signal relative to voxel-level signal. Easier to estimate $\rho$.
* For fixed $\delta_j$, larger $\psi_j$ means stronger spatial signal.

In simulated data, we choose parameters such that $\delta_j = 0.1, 0.5, 0.7$ and $\psi_j = 0.2, 0.5, 0.8$.
</div>


##

```{r}
#| echo=FALSE, message=FALSE, warning=FALSE, fig.width=12, fig.height=10, fig.align='center',
ebluedf <- readRDS(here("full-run", "out-eblue", "eblue-df.rds"))
oracledf <- readRDS(here("full-run", "out-oracle", "oracle-df.rds")) |>
              filter(method != "CA") |>
              mutate(method = as.factor("oracle"))
correct_specdf <- bind_rows(oracledf, ebluedf, ggdf) |>
          mutate(method = factor(method, levels = c("CA", "eblue", "oracle", "ReML")))
p <- create_boxplot(correct_specdf, "Comparing CA, ReML, EBLUE, and Oracle estimates")
print(p)
```


## Misspecification
We consider both temporal and spatial misspecification.

<div style="font-size: 24px;">

* Temporal misspecification: $A$ and $B_j$
   - $AR(2)$ with $\phi_1 = 0.4$ and $\phi_2 = 0.3$
   - Fractional Gaussian noise with $H = 0.7$
   - Pass raw signal through wavelet transformation to get 60 coefficients

</div>

. . .

<div style="font-size: 24px;">

* Spatial misspecification: $C_j$ [Castruccio et al. (2018)]
   - Anisotropic Matérn
   - Partition region $j$ into two subregions with centroids $v_{j1}^*, v_{j2}^*$.
   - Consider $R = \operatorname{diag}(1, 1.2, 1.5)$.
   - $[C_j]_{\ell\ell'} = \sum_{s=1}^2 w_{s\ell} w_{s\ell'} K(\big\lVert{R(v_{j\ell} - v_{j\ell'})}\big\rVert; \nu_j, \phi_j)$
   - $w_{sl} = (\big\lVert{v_{j\ell} - v_{js^*}}\big\rVert)^{-1}$ is the inverse distance to the centroid.

</div>

##

```{r}
#| echo=FALSE, message=FALSE, warning=FALSE, fig.width=12, fig.height=8, fig.align='center',
#| fig.cap="Temporal misspecification"

ggdf <- readRDS(here("full-run", "full-ar2-df.rds"))
p <- create_boxplot(ggdf, "Simulation results - AR(2) temporal covariance")
print(p)
```

##

```{r}
tabs1 <- filter(ggdf, method == "ReML") |>
         group_by(delta, psi, pair) |>
         summarize(mean = mean(value),
                   median = median(value),
                   mad = mean(abs(value - yintercept)),
                   rmse = sqrt(mean((value - yintercept)^2)),
                   sd = sd(value),
                   .groups = 'drop')

# Split the data into smaller parts
tabs1_r12 <- filter(tabs1, pair == "r12") |>
   select(-pair) |>
   mutate(across(-c(delta, psi), round, 3))
tabs1_r13 <- filter(tabs1, pair == "r13") |>
   select(-pair) |>
   mutate(across(-c(delta, psi), round, 3))
tabs1_r23 <- filter(tabs1, pair == "r23") |>
   select(-pair) |>
   mutate(across(-c(delta, psi), round, 3))
```

::: {.panel-tabset}

### Region 1 and 2

```{r}
kable(tabs1_r12, "html") |>
   kable_styling(font_size = 30)
```

### Region 1 and 3

```{r}
kable(tabs1_r13, "html") |> kable_styling(font_size = 30)
```

### Region 2 and 3

```{r}
kable(tabs1_r23, "html") |> kable_styling(font_size = 30)
```

:::

##

```{r}
#| echo=FALSE, message=FALSE, warning=FALSE, fig.width=12, fig.height=8, fig.align='center',
#| fig.cap="Temporal misspecification"

ggdf <- readRDS(here("full-run", "full-fgn-df.rds"))
p <- create_boxplot(ggdf, "Simulation results - fGn temporal covariance", 100)
print(p)
```

##

```{r}
tabs1 <- filter(ggdf, method == 'ReML') |>
         group_by(delta, psi, pair) |>
         summarize(mean = mean(value),
                   median = median(value),
                   mad = mean(abs(value - yintercept)),
                   rmse = sqrt(mean((value - yintercept)^2)),
                   sd = sd(value),
                   .groups = 'drop')

# Split the data into smaller parts
tabs1_r12 <- filter(tabs1, pair == "r12") |>
   select(-pair) |>
   mutate(across(-c(delta, psi), round, 3))
tabs1_r13 <- filter(tabs1, pair == "r13") |>
   select(-pair) |>
   mutate(across(-c(delta, psi), round, 3))
tabs1_r23 <- filter(tabs1, pair == "r23") |>
   select(-pair) |>
   mutate(across(-c(delta, psi), round, 3))
```

::: {.panel-tabset}

### Region 1 and 2

```{r}
kable(tabs1_r12, "html") |>
   kable_styling(font_size = 30)
```

### Region 1 and 3

```{r}
kable(tabs1_r13, "html") |> kable_styling(font_size = 30)
```

### Region 2 and 3

```{r}
kable(tabs1_r23, "html") |> kable_styling(font_size = 30)
```

:::

##

```{r}
#| echo=FALSE, message=FALSE, warning=FALSE, fig.width=12, fig.height=8, fig.align='center',
#| fig.cap="Spatial misspecification"

ggdf <- readRDS(here("full-run", "full-anisotropic-df.rds"))
p <- create_boxplot(ggdf, "Simulation results - Anisotropic spatial covariance")
print(p)
```

##

```{r}
tabs1 <- filter(ggdf, method == 'ReML') |>
         group_by(delta, psi, pair) |>
         summarize(mean = mean(value),
                   median = median(value),
                   mad = mean(abs(value - yintercept)),
                   rmse = sqrt(mean((value - yintercept)^2)),
                   sd = sd(value),
                   .groups = 'drop')

# Split the data into smaller parts
tabs1_r12 <- filter(tabs1, pair == "r12") |>
   select(-pair) |>
   mutate(across(-c(delta, psi), round, 3))
tabs1_r13 <- filter(tabs1, pair == "r13") |>
   select(-pair) |>
   mutate(across(-c(delta, psi), round, 3))
tabs1_r23 <- filter(tabs1, pair == "r23") |>
   select(-pair) |>
   mutate(across(-c(delta, psi), round, 3))
```

::: {.panel-tabset}

### Region 1 and 2

```{r}
kable(tabs1_r12, "html") |>
   kable_styling(font_size = 30)
```

### Region 1 and 3

```{r}
kable(tabs1_r13, "html") |> kable_styling(font_size = 30)
```

### Region 2 and 3

```{r}
kable(tabs1_r23, "html") |> kable_styling(font_size = 30)
```

:::
